Supporting information: Copulas reveal complex and informative dependencies propagating throughout ecology

### 2019-05-15 Paper.Rmd
small chnages suggested by Shya, Lawrence, Van
cedar creek data refs added

Datasets included environmental, species-trait, phenological, 
population, community, and ecosystem functioning data (Table \ref{tab_data_info}). Copula structure describing 
the relationship between atmospheric weather variables such as rainfall or wind speed 
as measured in multiple locations has been 
examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database [RaCA; @Wills2014].
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
body mass (grams) for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
Survey (RIS) suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years. 
Our second population dataset comprised average annual plankton abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. 
Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` plots, each 9m by 9m, as sampled in 
the years 1996-2000 and 2007, each year analyzed separately [@Tilman2001; @Tilman2006]. 
Shannon's index was computed from plant species percent cover data. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. See Appendix \ref{SM-Data_details} for additional data details.

@@

# Background, concepts and methods for testing the observational hypothesis \label{Methods}

@@

# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{1pt}
\noindent

>>>>

Datasets included environmental, species-trait, phenological, 
population, community, and ecosystem functioning data (Table \ref{tab_data_info}). Copula structure describing 
the relationship between atmospheric weather variables such as rainfall or wind speed 
as measured in multiple locations has been 
examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database [RaCA; @Wills2014].
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
body mass (grams) for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
Survey (RIS) suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years. 
Our second population dataset comprised average annual plankton abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. 
Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
plant aboveground biomass [@ccdataple120] and Shannon's diversity index 
for `r nrow(d_cc[[5]])` plots[@ccdataplotse120], each 9m by 9m, as sampled in 
the years 1996-2000 and 2007, each year analyzed separately [@Tilman2001; @Tilman2006]. 
Shannon's index was computed from plant species percent cover data[@ccdatapce120]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. See Appendix \ref{SM-Data_details} for additional data details.

@@

# Background, concepts and methods for testing the observational hypothesis\label{Methods}

@@

# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{1pt}
\noindent
\sloppy

### 2019-05-15 SupportingInformation.Rmd
cedar creek data refs added

\textbf{Community-level data, Cedar Creek.} Above-ground biomass and percent cover data were 
obtained from biodiversity experiment E120 conducted at the Cedar Creek Ecosystem Science Reserve 
by Tilman *et al*. In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
were seeded with 1, 2, 4, 8, or 16 perennials, respectively, chosen randomly from a pool of 
18 species (4 species, each, of C4 grasses, C3 grasses, legumes, non-legume forbs; and 2 species of woody
plants). All plots received, in total, 10 $gm^{-2}$ of pure live seed in May 1994 and 5 $gm^{-2}$ in May, 1995 
with seed mass divided equally among species. Treatments were maintained by weeding manually 3-4 times a year 
and sometimes selective herbicides were used. We downloaded the data on July, 2018 from the Cedar Creek website \url{http://www.cedarcreek.umn.edu/research/experiments/e120} and 
analyzed those data for 6 years 
(1996-2000, 2007) for which both biomass and percent cover data were available. Plots were annually 
sampled by clipping 
different locations each year. For each available year, we analyzed above-ground biomass ($gm^{-2}$) 
and Shannon's biodiversity index (computed using percent cover data) averaged over the 
sampled strips/quadrats taken from each 
of 168 plots.

>>>>

\textbf{Community-level data, Cedar Creek.} Above-ground biomass [@ccdataple120] and percent cover data [@ccdatapce120] were 
obtained from biodiversity experiment E120 conducted at the Cedar Creek Ecosystem Science Reserve 
by Tilman *et al*. In 1994, 168 plots [@ccdataplotse120] (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
were seeded with 1, 2, 4, 8, or 16 perennials, respectively, chosen randomly from a pool of 
18 species (4 species, each, of C4 grasses, C3 grasses, legumes, non-legume forbs; and 2 species of woody
plants). All plots received, in total, 10 $gm^{-2}$ of pure live seed in May 1994 and 5 $gm^{-2}$ in May, 1995 
with seed mass divided equally among species. Treatments were maintained by weeding manually 3-4 times a year 
and sometimes selective herbicides were used. We downloaded the data on July, 2018 from the Cedar Creek website \url{http://www.cedarcreek.umn.edu/research/experiments/e120} and 
analyzed those data for 6 years 
(1996-2000, 2007) for which both biomass and percent cover data were available. Plots were annually 
sampled by clipping 
different locations each year. For each available year, we analyzed above-ground biomass ($gm^{-2}$) 
and Shannon's biodiversity index (computed using percent cover data) averaged over the 
sampled strips/quadrats taken from each 
of 168 plots.
